geochemistry

### 2025-01-29 geochemistry.qmd
Boxplots of all relevant elements. Scatterplots.

library(readr)
library(dplyr)
library(janitor)
library(tidyr)
library(gt)
library(ggplot2)

@@

@@

pxrf_average <- pxrf_analysis |> 
  select(-Sample_ID, -pXRF_ID) |> 
  mutate(across(-`Formation/Type`, ~ as.numeric(.))) |>    # Convert valid numbers
  mutate(across(where(is.numeric), ~ ifelse(is.na(.), 0, .))) |>  # Replace NAs with 0
  group_by(`Formation/Type`) |> 
  summarise(across(where(is.numeric), ~ round(mean(.), 3)))
  pivot_wider(names_from = `Formation/Type`,
              values_from = c("MgO", "Al2O3", "SiO2", "P2O5", "S", "Cl", "K2O", "CaO", "TiO2", "Fe2O3", "Cu",
         "Sr", "Zr"))

>>>>

library(readr)
library(dplyr)
library(janitor)
library(tidyr)
library(gt)
library(ggplot2)
library(cowplot)

@@

```{r}
#| label: boxplots

# Boxplot data preparation
boxplots_pxrf <- pxrf_analysis |> 
  select(-Sample_ID, -pXRF_ID) |> 
  mutate(across(-`Formation/Type`, ~ as.numeric(.)))
  #pivot_longer(cols = c("MgO", "Al2O3", "SiO2", "P2O5", "S", "Cl", "K2O", "CaO", "TiO2", "Fe2O3", "Cu",
        # "Sr", "Zr"), names_to = "Variable", values_to = "Value")
boxplots_pxrf <- boxplots_pxrf |> 
  mutate(`Formation/Type` = factor(`Formation/Type`, levels = c("TW", "LJW", "MJE", "UJW", "UJE", "CPT", "UCA", "T1", "T2",
                                                                "T3", "T4", "T5", "T6", "T7", "T8", "T9", "T10", "T11")))
## MgO
count_data_mgo <- boxplots_pxrf |> # Count number of observations per group where Value > 0
  select(`Formation/Type`, MgO) |> 
  group_by(`Formation/Type`) |> 
  filter(!is.na(MgO)) |> 
  summarise(n = n(), .groups = "drop")

b1 <- boxplots_pxrf |> 
  ggplot(aes(x = `Formation/Type`, y = MgO)) +
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 18) +
  geom_text(data = count_data_mgo, aes(x = `Formation/Type`, y = 0.5, label = n), size = 3) +
  theme_minimal()

## Al2O3
count_data_al2o3 <- boxplots_pxrf |> # Count number of observations per group where Value > 0
  select(`Formation/Type`, Al2O3) |> 
  group_by(`Formation/Type`) |> 
  filter(!is.na(Al2O3)) |> 
  summarise(n = n(), .groups = "drop")

b2 <- boxplots_pxrf |> 
  ggplot(aes(x = `Formation/Type`, y = Al2O3)) +
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 18) +
  geom_text(data = count_data_al2o3, aes(x = `Formation/Type`, y = 0, label = n), size = 3) +
  theme_minimal()

## SiO2
count_data_sio2 <- boxplots_pxrf |> # Count number of observations per group where Value > 0
  select(`Formation/Type`, SiO2) |> 
  group_by(`Formation/Type`) |> 
  filter(!is.na(SiO2)) |> 
  summarise(n = n(), .groups = "drop")

b3 <- boxplots_pxrf |> 
  ggplot(aes(x = `Formation/Type`, y = SiO2)) +
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 18) +
  geom_text(data = count_data_sio2, aes(x = `Formation/Type`, y = 40, label = n), size = 3) +
  theme_minimal()

## P2O5
count_data_p2o5 <- boxplots_pxrf |> # Count number of observations per group where Value > 0
  select(`Formation/Type`, P2O5) |> 
  group_by(`Formation/Type`) |> 
  filter(!is.na(P2O5)) |> 
  summarise(n = n(), .groups = "drop")

b4 <- boxplots_pxrf |> 
  ggplot(aes(x = `Formation/Type`, y = P2O5)) +
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 18) +
  geom_text(data = count_data_p2o5, aes(x = `Formation/Type`, y = 0, label = n), size = 3) +
  theme_minimal()

## S
count_data_s <- boxplots_pxrf |> # Count number of observations per group where Value > 0
  select(`Formation/Type`, S) |> 
  group_by(`Formation/Type`) |> 
  filter(!is.na(S)) |> 
  summarise(n = n(), .groups = "drop")

b5 <- boxplots_pxrf |> 
  ggplot(aes(x = `Formation/Type`, y = S)) +
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 18) +
  geom_text(data = count_data_s, aes(x = `Formation/Type`, y = -0.1, label = n), size = 3) +
  theme_minimal()

## Cl
count_data_cl <- boxplots_pxrf |> # Count number of observations per group where Value > 0
  select(`Formation/Type`, Cl) |> 
  group_by(`Formation/Type`) |> 
  filter(!is.na(Cl)) |> 
  summarise(n = n(), .groups = "drop")

b6 <- boxplots_pxrf |> 
  ggplot(aes(x = `Formation/Type`, y = Cl)) +
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 18) +
  geom_text(data = count_data_cl, aes(x = `Formation/Type`, y = -0.05, label = n), size = 3) +
  coord_cartesian(ylim = c(-0.05, 0.75)) +
  theme_minimal()

## K2O
count_data_k2o <- boxplots_pxrf |> # Count number of observations per group where Value > 0
  select(`Formation/Type`, K2O) |> 
  group_by(`Formation/Type`) |> 
  filter(!is.na(K2O)) |> 
  summarise(n = n(), .groups = "drop")

b7 <- boxplots_pxrf |> 
  ggplot(aes(x = `Formation/Type`, y = K2O)) +
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 18) +
  geom_text(data = count_data_k2o, aes(x = `Formation/Type`, y = -0.05, label = n), size = 3) +
  theme_minimal()

## CaO
count_data_cao <- boxplots_pxrf |> # Count number of observations per group where Value > 0
  select(`Formation/Type`, CaO) |> 
  group_by(`Formation/Type`) |> 
  filter(!is.na(CaO)) |> 
  summarise(n = n(), .groups = "drop")

b8 <- boxplots_pxrf |> 
  ggplot(aes(x = `Formation/Type`, y = CaO)) +
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 18) +
  geom_text(data = count_data_cao, aes(x = `Formation/Type`, y = -0.3, label = n), size = 3) +
  theme_minimal()

## TiO2
count_data_tio2 <- boxplots_pxrf |> # Count number of observations per group where Value > 0
  select(`Formation/Type`, TiO2) |> 
  group_by(`Formation/Type`) |> 
  filter(!is.na(TiO2)) |> 
  summarise(n = n(), .groups = "drop")

b9 <- boxplots_pxrf |> 
  ggplot(aes(x = `Formation/Type`, y = TiO2)) +
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 18) +
  geom_text(data = count_data_tio2, aes(x = `Formation/Type`, y = -0.02, label = n), size = 3) +
  theme_minimal()

## Fe2O3
count_data_fe2o3 <- boxplots_pxrf |> # Count number of observations per group where Value > 0
  select(`Formation/Type`, Fe2O3) |> 
  group_by(`Formation/Type`) |> 
  filter(!is.na(Fe2O3)) |> 
  summarise(n = n(), .groups = "drop")

b10 <- boxplots_pxrf |> 
  ggplot(aes(x = `Formation/Type`, y = Fe2O3)) +
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 18) +
  geom_text(data = count_data_fe2o3, aes(x = `Formation/Type`, y = 0, label = n), size = 3) +
  theme_minimal()

## Cu
count_data_cu <- boxplots_pxrf |> # Count number of observations per group where Value > 0
  select(`Formation/Type`, Cu) |> 
  group_by(`Formation/Type`) |> 
  filter(!is.na(Cu)) |> 
  summarise(n = n(), .groups = "drop")

b11 <- boxplots_pxrf |> 
  ggplot(aes(x = `Formation/Type`, y = Cu)) +
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 18) +
  geom_text(data = count_data_cu, aes(x = `Formation/Type`, y = 0, label = n), size = 3) +
  theme_minimal()

## Sr
count_data_sr <- boxplots_pxrf |> # Count number of observations per group where Value > 0
  select(`Formation/Type`, Sr) |> 
  group_by(`Formation/Type`) |> 
  filter(!is.na(Sr)) |> 
  summarise(n = n(), .groups = "drop")

b12 <- boxplots_pxrf |> 
  ggplot(aes(x = `Formation/Type`, y = Sr)) +
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 18) +
  geom_text(data = count_data_sr, aes(x = `Formation/Type`, y = -0.005, label = n), size = 3) +
  theme_minimal()

## Zr
count_data_zr <- boxplots_pxrf |> # Count number of observations per group where Value > 0
  select(`Formation/Type`, Zr) |> 
  group_by(`Formation/Type`) |> 
  filter(!is.na(Zr)) |> 
  summarise(n = n(), .groups = "drop")

b13 <- boxplots_pxrf |> 
  ggplot(aes(x = `Formation/Type`, y = Zr)) +
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 18) +
  geom_text(data = count_data_zr, aes(x = ` `, y = 0, label = n), size = 3) +
  theme_minimal()

boxplots_pxrf_all <- pxrf_analysis |> 
  select(-Sample_ID, -pXRF_ID) |> 
  mutate(across(-`Formation/Type`, ~ as.numeric(.))) |> 
  pivot_longer(cols = c("MgO", "Al2O3", "SiO2", "P2O5", "S", "Cl", "K2O", "CaO", "TiO2", "Fe2O3", "Cu",
        "Sr", "Zr"), names_to = "Variable", values_to = "Value")

boxplots_pxrf_all |> 
  ggplot(aes(x = `Formation/Type`, y = Value)) +
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 18) +
  facet_wrap(~Variable, scales = "free") +
  #geom_text(data = count_data_zr, aes(x = ` `, y = 0, label = n), size = 3) +
  theme_minimal()

```{r}

plot_grid(b1, b2, b3, b4, b5, b6, b7, b8, b9, b10, b11, b12, b13, labels = c("a","b","c","d","e","f","g","h","i","j","k","l","m"))

```{r}

# Define the list of elements
elements <- c("MgO", "Al2O3", "SiO2", "P2O5", "S", "Cl", "K2O", 
              "CaO", "TiO2", "Fe2O3", "Cu", "Sr", "Zr")

# Generate all possible combinations of 2 elements
combinations <- combn(elements, 2, simplify = FALSE)

comb_df <- as.data.frame(do.call(rbind, combinations))
colnames(comb_df) <- c("Element1", "Element2")

# Print all combinations
print(combinations)

ggplot(boxplots_pxrf, aes(x = Sr, y = Zr, color = `Formation/Type`)) + 
  geom_point(shape = 21, size = 4, fill = "white", stroke = 1.2) +  # Empty circles with border
  stat_ellipse(level = 0.95, size = 1.2, linetype = "dashed") +
  theme_minimal()

@@

pxrf_average <- pxrf_analysis |> 
  select(-Sample_ID, -pXRF_ID) |> 
  mutate(across(-`Formation/Type`, ~ as.numeric(.))) |>    # Convert valid numbers
  mutate(across(where(is.numeric), ~ ifelse(is.na(.), 0, .))) |>  # Replace NAs with 0
  group_by(`Formation/Type`) |> 
  summarise(across(where(is.numeric), ~ round(mean(.), 4)))
  pivot_wider(names_from = `Formation/Type`,
              values_from = c("MgO", "Al2O3", "SiO2", "P2O5", "S", "Cl", "K2O", "CaO", "TiO2", "Fe2O3", "Cu",
         "Sr", "Zr"))
